Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-01-24 manuscript/supplementaryRW.Rmd
Remove duplicate heading for Formind
correct a small spelling error in Formind suplementary report
update report from TreeMig in supplementary

We did not use any topograpy data.

@@

@@

@@

What climate data have you used, and how have you created your model
climate input parameters? I used the monthly time series of tas and
precipitation,additionally slope and aspect from the topo- data set, and
soil quality from the quality data set. slasp= 2*cos(aspect* pi/180)
\*min(1, slope/60)

@@

The soil quality I translated into AWC (of  upmost m soil, "bucketsize")
by Bucketsize = 10 + (soil_quality[-1) /4  \* 25 , to covert he range
from 10 cm to 35cm   With these values I calculated the yearly time
series of the bioclimatic variables DDsum (day degree sum above 5.5°C) ,
minWinterTemp (lowest mean monthly temperature) and Drought stress
according to a bucket model, that estimates a monthly PET following
[Thorntwaithe, 1957],  and calculates the monthly soil water content
   based on soil water content in the previous month, precipitation,
interception, and bucketsize   [Bugmann , 1994] . Drought stress is then
given by 1- the ratio between demand (PET-interception) and the supply
(a function of the soil water content). Extending the climate data I
used the monthly data as given, calculated the bioclimate, and extended
since the end of the climate data to get 1000 years, by sampling from
the given data. I used always the same sequence of sampling years in all
simulations and sites. (-\> better comparability of experiments)

@@

@@

How have you defined the species composition in your model simulation?
(e.g., species levels, PFTs, etc.)? Please describe the species
categories used. Species level, 30 species (see list of species I sent).

@@

How have you defined the simulation area for your model (e.g., size and
other relevant site aspects)? A grid of 15 \* 15 grid cells, each grid
cell with 200m side length.

@@

What are the relevant aspects of the dispersal (if any) and seed input
parameters in your model? Seed dispersal is described by species
specific, height dependent seed production. The seeds are then
transported according to a double negative exponential, species specific
kernel, and enter the seed bank in the sink cells, from which seedlings
germinate depending on light and other environmental conditions. At the
start of the simulation, for 10 years seeds of all species are available
in all grid cells, from then on, the normal seed dispersal starts.

There are 5 different experiments related to dispersal NoDisp:  the
dispersal kernels are so narrow, that the huge majority of seeds stays
in the source grid cell. I.e. the spatial variability (see below) only
arises from the sampling of the light classes. But there is still the
feedback between the mother trees and the seedlings. StochDisp:
stochastic dispersal. The seeds are distributed randomly, whereby the
kernel gives the probability of the seeds  landing in a certain distance
DetDisp: deterministic dispersal. The kernel determines the share of the
seeds in the source cell landing in a sink cell in a given distance to
the source cell. FBirthAll: a fixed ingrowth above 1.37m of 20
trees/833m\^2  and year, same for all species, but reduced by
environmental drivers and shade.  This is analogous to the approach of
ForClim,  i.e. ignores the feedback  between mother trees and young
trees and the youth development FBirthIndiv: as FBirthAll, but the fixed
ingrowth is species-specific  (between 4 and 54 trees/833m\^2 )

@@

What is the simulation length, and how have you determined the
simulation time until the equilibrium? 1000 years, experience and check
by eye-ball of the simulated basal areas. Variability: I have employed
three different aspects of the model  to account for spatial and
temporal variability and to get the required 200 patches of 1ha.

1)  Basic spatial variability due to the combination of demographic
    stochasticity and short-range spatial interactions. In contrast to
    gap models, TreeMig is based on the concept of frequency
    distributions of tree densities classes in different tree heights on
    ca. 1/12 ha (833m\^2) patches, and the resulting frequency
    distributions in discrete light. During the simulation, in each
    height class the frequencies of the different light classes are
    recorded, as well as the ingrowth of each species in the
    height-light class. By sampling 12 times from the height-light class
    ingrowths according to the height-light class frequency
    distribution, and summing up, the stochastic ingrowth for one ha is
    calculated, for each grid cell.  Then the height class boundaries
    are translated into DBH and the ingrowths in the height classes
    linearly interpolated to 7cm and 10cm.

2)  Spatial smoothing and variability by the formulation of dispersal.
    The different dispersal strategies, see above. With NoDisp, solely
    the basic spatial variability is accounted for. With DetDisp, I
    expect the dispersal to smooth out the variability. StochDisp, in
    turn is expected to create additional variability by stochastic
    founder processes.

3)  Temporal variability by sampling from different decades of the last
    100 years of the  simulation, i.e. years 900-1000

#### Regeneration parameters

How does environment affect regeneration in the model? Seed production:
depends on tree size and thus indirectly on the growth, that  depends on
DDsum,  drought stress, nutrients and light and on adult tree number and
thus on survival  that depends on DDsum, drought stress, nutrients and
light

Seed dispersal:  independent on environment Seed bank dynamics : 
independent on environment Germination and seedling survival: Depends on
drought, browsing, winter temperature and DDSum and light (last 3 are
step functions) Sapling dynamics (up to 1.37): like adult growth and
survival, i.e. depends on DDsum, drought stress and light but light
dependence with a specific parameter values, namely the klights
parameter instead of klighta

#### more simulations per site

If you have run more than one simulation per site, please describe the
differences of the approaches used. I used the following 15
combinationsExperiments (exp_txpe) NoDisp, StochDisp, DetDisp,
FBirthAll, FBirthIndiv), each with the variability setups (comb_type)
200space1time: 200 grid cells at a single randomly chosen decade (
50space4time: 50 grid cells at 4 randomly chosen decades, or
20space10time: 20 grid cells at all 10 decades Always the same random
(nested) sets of decades and grid cells were used Fixed ingrowth values
(for the FBirthAll and FBirthIndiv experiments ) were calculated
beforehand in a  set of simulations with single species under optimal
conditions, without interspecific, but under intraspecific competition.
The ingrowth into the height class above 1.37m per species was averaged
over the last 100  years of the 1000 year simulation and used in the
FBirthIndiv simulations. For the FBirthAll simulations the mean of the
single species ingrowths' (20) was used.

The "standard" combination is DetDisp with 20 grid cells at all 10
decades Please describe how have you calculated each of the variables in
the output data. During the simulation,  in each height class the
frequencies of the different light classes are recorded, as well as the
ingrowth numbers of each species in the height-light class. By sampling
12 times from the height-light class ingrowth numbers according to the
height-light class  frequency distribution, and summing up, the
stochastic ingrowth number for one ha is calculated, for each grid
cell.  Then the height class boundaries are translated into DBH and the
ingrowths in the height classes linearly interpolated to 7cm and 10cm
based.  The ingrowth basal areas into the height classes are obtained by
multiplying the interpolated ingrowth numbers with the basal areas at 
7cm and 10cm, i.e. r.trees \* Pi *(0.07/2)\^2  and r.trees* Pi\* 
(0.1/2)\^2 .

The species specific basal area (at 1.37m) sums are a standard output of
TreeMig, given by multiplying the state variables (number of trees per
grid cell , height class, and species) with  the species specific basal
area of each height class. Files TreeMigRes_all_data.RData: R RData-file
with all experiments (exp_type) and variability combinations (comb_type)
TreeMigRes_Std_data.RData: R RData-file with   exp_type=detDisp) and
variability combinations (comb_type=20space10time)

@@

>>>>

We did not use any topography data.

@@

### Authors
Heike Lischke

@@

The model TreeMig (see treemig.wsl.ch) was originally based on the ForClim model, then aggregated by a distribution based approach (Lischke et al. 1998), and extended to a forest landscape model by including seed production, seed dispersal, seed bank dynamics and seedling establishment (Lischke and Loffler 2006, Lischke et al. 2006).

@@

As climatic input, the monthly time series of mean temperature and precipitation sum were used.  With these climatic variables and additional data about slope, aspect and soil field capacity, yearly time series of the following bioclimatic variables were calculated:  (1) DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest mean monthly temperature, (3) DrStr, drought stress according to a bucket model  that estimates a monthly PET following (Thornthwaite and Mather 1957), using temperature, latitude, slope and aspect as input,  and calculates the monthly soil water content  based on soil water content in the previous month, precipitation, interception, and field capacity (“bucketsize”)   (Bugmann 1994, Fischlin et al. 1995). Drought stress is then given by 1- the ratio between demand (PET-interception) and the supply, a function of the soil water content.
To get climate data also for the spin-up, bioclimate was calculated from the monthly data   as given, and then extended by sampling from the given data from the end of the climate data to get 1000 years. The same sequence of sampling years was used in all simulations and sites.

@@

The soil bucketsize (available water capacity of the upmost 1m soil layer) required for the drought stress calculation, was derived from the soil quality data set, by
Bucketsize = 10 + (soil_quality-1) /4  * 25 , to covert the range from 10 cm to 35cm

@@

Slope and aspect were taken from the topographic data set and converted to a common variable entering the PET calculation, by
slasp= 2*cos(aspect *pi/180) *min(1, slope/60)

@@

Thirty Central European tree species were simulated, including the standard species of the project .  Only Betula pubescens was not simulated, because we did not have a parametrization for this species at hand. Species: Abies alba, Larix decidua, Picea abies, Pinus cembra, Pinus montana, Pinus sylvestris, Taxus baccata, Acer campestre, Acer platanoides, Acer pseudoplatanus, Alnus glutinosa, Alnus incana, Alnus viridis, Betula pendula, Carpinus betulus, Castanea sativa, Corylus avellana, Fagus sylvatica, Fraxinus excelsior, Populus nigra, Populus tremula, Quercus petraea, Quercus pubescens, Quercus robur, Salix alba, Sorbus aria, Sorbus aucuparia, Tilia cordata, Tilia platyphyllos, Ulmus glabra

@@

The simulations were carried out on a grid of 15 x 15 cells, each with 200 m side length.

@@

The seeds produced by the mother trees are transported from the source cell according to a double negative exponential, species specific that determines the share of the seeds in the source cell landing in a sink cell in a given distance to the source cell.  In the sink cells, the seeds enter the seed bank . At the start of the simulation for 10 years seeds of all species are available in all grid cells, from then on, the normal seed dispersal starts.

#### Regeneration processes

TreeMig simulates the full feedback from seed production to ingrowth. Some of the involved processes depend on environment.
(1) Overall seed production in TreeMig is described by multiplying a species specific seed production per tree that depends on tree size and thus indirectly on the growth, which depends on DDsum, drought stress, nutrients and light and the adult tree number that depends on survival  that depends on DDsum, drought stress, nutrients and light.
(2) Seed dispersal is independent of environment. 
(3) Seed bank dynamics – also environment independent -   is given by seed input, seed mortality, and germination.  Additionally, the number of seeds S of a species in the seed bank of this species (SB) is limited to a carrying capacity C of 1000, by           
SB = C + (SB - C) * Exp(-S/C).  
(4) Germination and seedling survival depend on browsing, drought, winter temperature, DDSum and light. 
(5) The saplings up to 1.37 die and grow like adult trees, depending on DDsum, drought stress and light. Only the parameter value of the light dependence differs from that of the adult trees.  
The regeneration processes were simulated annually in the order 1. germination, 2. germinated seeds leave the seed bank, 3. mortality of seeds in the seed bank, 4. seed entry by seed dispersal. The ingrowth numbers were very sensitive to this order. Simulations with another sequence, in which the seeds in the seed bank died first and then  the  remaining ones germinated,  resulted in about half the ingrowth.

@@

The simulations were started on bare ground and run for 1000 years, when at all sites the equilibrium was reached.

#### More simulations per site

TreeMig simulates directly the mean dynamics of entire stands, but based on the  variability of light conditions within the stand   [distribution based approach, \Lischke, 1998, aggregation. To come up with the required 200 random patches of 1 ha, temporal and spatial variability intrinsic to TreeMig were combined.  Temporal variability was given by sampling from each decade of the last 100 years of the simulation, i.e. years 900-1000. For each of theses 10 decades, 20 grid cells were sampled from the simulated grid, reflecting spatial variability, results from demographic stochasticity and short-range spatial interactions

@@

In contrast to gap models, TreeMig is based on the concept of frequency distributions of tree densities classes in different tree heights on ca. 1/12 ha (833m^2) patches, and the resulting frequency distributions in discrete light. During the simulation, in each height class the frequencies of the different light classes were recorded, as well as the ingrowth of each species in the height-light class. By sampling 12 times from the height-light class ingrowths according to the height-light class frequency distribution and summing up, the stochastic ingrowths into all TreeMig height classes for one ha were calculated for each grid cell.  Then the height class boundaries were translated into DBH and the ingrowths into the height classes linearly interpolated to 7cm and 10cm. 
The ingrowth basal areas into the height classes were obtained by multiplying the interpolated ingrowth numbers with the basal areas at  7cm and 10cm, i.e. r.trees * Pi*  (0.07/2)2  and r.trees * Pi*  (0.1/2)2 .
The species specific basal area (at 1.37m) sums are a standard output of TreeMig, given by multiplying the state variables (number of trees per grid cell, height class, and species) with  the species specific basal area of each height class. 
To account for the ingrown trees dying within a decade, as an approximation the survival in each year to the power of 5 was used. Thereby, it was assumed that the trees have to survive in average 5 years in each decade, and that the current mortality is valid for these 5 years.

•	Bugmann, H. 1994. On the ecology of mountainous forests in a changing climate: A simulation study. Dissertation. Swiss Federal Institute of Technology, Zurich.
•	Fischlin, A., H. Bugmann, and D. Gyalistras. 1995. Sensitivity of a forest ecosystem model to climate parametrization schemes. Env. Poll. 87:267-282.
•	Lischke, H., and T. J. Loffler. 2006. Intra-specific density dependence is required to maintain species diversity in spatio-temporal forest simulations with reproduction. Ecological Modelling 198:341-361.
•	Lischke, H., T. J. Loffler, and A. Fischlin. 1998. Aggregation of individual trees and patches in forest succession models: Capturing variability with height structured, random, spatial distributions. Theor Popul Biol 54:213-226.
•	Lischke, H., N. E. Zimmermann, J. Bolliger, S. Rickebusch, and T. J. Loffler. 2006. TreeMig: A forest-landscape model for simulating spatio-temporal patterns from stand to landscape scale. Ecological Modelling 199:409-420.
•	Thornthwaite, C. W., and J. R. Mather. 1957. The water balance. Publications in Climatology VIII:1-69.
